---
title: "Lab 5 markdown"
author: "Victoria Wallace"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Lab 5 Objectives

- Learn to parse dates with 'lubridate'
- Get coutns of observations with count()
- Uncount() rows
- One and two-sample t-tests
- Create a heat map with geom_tile()


```{r, include = FALSE}
# Read in data and attach packages.
library(tidyverse)
library(here)
library(janitor)

lobster_abundance <- read_csv(here::here("data", "lobster_abundance.csv"),
                              na = "-99999") %>% 
  clean_names()
```
Use tidyr::uncount() to convert lobster data from frequency format to case format.
```{r}
lobster_tidy <-  lobster_abundance %>% 
  uncount(lobster_count)
```

#### Exploratory data visualization

We're only going to consider 'site' as our variable of interest.
```{r}
# Jitterplot
ggplot(lobster_tidy, aes(x = site,
                         y = size_mm)) +
  geom_jitter(aes(color = site),
              alpha = 0.5,
              width = 0.2)

# Histograms
ggplot(lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free") # Scales = "free" is good for when you want to compare distrubutions, not relative differences in magnitude

ggplot(lobster_tidy, aes(sample = size_mm)) +
  geom_qq() +
  facet_wrap(~site)
```
#### Convert the date column to class "Date"

We'll use the lubridate package to convert to date format, and then to parse month & year.
```{r}
lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
  )
```

Now, let's parse year and month using lubridate::month() and lubridate::year().

```{r}
lobster_parse_date <-  lobster_date %>% 
  mutate(obs_month = lubridate::month(date_new, label = TRUE),
         obs_year = lubridate::year(date_new))
  
```

Now, let's find counts of observed lobsters based on different hierarchical groupings:

First, let's count lobsters by year and month:
```{r}
lobster_ym <- lobster_parse_date %>% 
  dplyr::count(obs_year, obs_month)

lobster_year <- lobster_parse_date %>% 
  count(obs_year)
```

I want to know how many lobsters were observed at each site.

```{r}
lobster_site <- lobster_parse_date %>% 
  count(site)
```

Having data in tidy formats gives the coder the freedom to form counts that are most important for data analysis and visualization.

#### Reporting summary statistics in a table

I'll use group_by() + n(). 

```{r}
lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    lobster_number = n(),
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE))
```

#### Find confidence intervals

Use t.test() to find confidence intervals for one sample and perform t-tests to compare means of two samples.

```{r}
ivee_lobsters <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

# If you want to convert a column in a data frame to a vector, use the pull() function.

ivee_ttest <-  t.test(ivee_lobsters)
ivee_ttest
```

#### 2-sample t-test to compare means

Is there a significant difference between two samples (do the populations have different means)?

- I've done the necessary exploratory analyses to determine that a 2-sample t-test for means comparison is appropriate. 
- The data look to be distributed normally, so comparing means makes sense.


```{r}
napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mohk_napl_ttest <- t.test(mohk_sample, napl_sample)

mohk_napl_ttest
```

t-statistic = 19.849

The difference in means is not equal to 0. But we want to know if there if evidence to support the hypothesis that these samples are drawn from populations that don't have the same mean.

Looking at the p-value, it is much more likely that the lobster populations have different mean lengths.

There is a significant difference in lobster lengths between Naples and Mohawk Reef (t(df) = statistic, p < 0.001, alpha = 0.05).

```{r}
lobster_mn <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

mn_ttest2 <- t.test(size_mm ~ site, data = lobster_mn)

mn_ttest2$p.value
```

#### How to use in-line referencing

The p-value is `r mn_ttest2$p.value`

#### Now, a heatmap

Used to show two discrete variables

```{r}
lobster_ys <- lobster_parse_date %>% 
  count(obs_year, site)

ggplot(data = lobster_ys,
       aes(x = obs_year,
           y = site)) +
  geom_tile(aes(fill = n))
```



